---
title: "Lab 6"
author: "Patrick Pelegri-O'Day"
date: "11/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library (tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Example of a rank-based test

We'll make our own samples, using a pseudorandom generator.

```{r}
set.seed(1414) # psuedorandom generator - everyone who uses this code will have the same random sample
gp_1 <- sample.int(20, size = 15, replace = TRUE) # gives a range from 0 to 20, sample size = 15, once the value of an observation is used, that number gets put back in the possible pool of numbers that can be chosen from, so a number can be used multiple times

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

```{r}
hist(gp_1)

hist(gp_2)
```

Try a t-test

```{r}
t.test(gp_1, gp_2)
```

This p-value means that if these two random samples were drawn from populations with the same sample mean, there is a probability of 19.8% that we would take two random samples that have means at least as different as the sample means we found by random chance.

As a result we would retain (fail to reject) the null hypothesis. (People get weirdly upset if you say "accept" the null.) There is no significant difference in means between group 1 and group 2.

